Generalized Tractability

### 2023-03-10 GeneralizedTractability_230310.tex
Fred M2 2023 Mac Updates

Certain kinds of common notions of tractability are 
\begin{equation*}
    \begin{array}{r@{\qquad}c@{\qquad}c}
    & \comp(\varepsilon, d)\\
    \toprule
    \text{strong polynomial algebraic  tractability} & \Theta(\varepsilon^{-p})\\
    \text{polynomial algebraic tractability} & \Theta(\varepsilon^{-p}d^{q})\\
    \text{strong polynomial exponential tractability} &  \Theta\bigl([\log(1 + \varepsilon^{-1})]^p\bigr)\\
    \text{polynomial exponential tractability} & \Theta\bigl([\log(1 + \varepsilon^{-1})]^p  d^{q}\bigr)
    \end{array}
\end{equation*}

@@

Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty) \times [0,\infty) \rightarrow [1,\infty),
\] 
which is non-decreasing in all variables.

Also, {we assume that, for any choice of $p\in[0,\infty)$, $q\in [0,\infty)$, and $\tau\in (1,\infty)$,
\[
[T(\varepsilon^{-1},d,p,q)]^\tau \le K_{p,q,\tau} T(\varepsilon^{-1},d,\tau p,\tau q),
\]
where $K_{p,q,\tau}$ is a term that may depend on $p$, $q$, and $\tau$, but is independent of $\varepsilon$ and $d$.

Furthermore, we assume that, for any choice of $d\in\naturals$ and $p,q\in [0,\infty)$, 
the quantity $T(0,d,p,q):=\liminf_{\varepsilon\to\infty}T(\varepsilon^{-1},d,p,q)$ exists.

A problem is $T$-tractable with parameters $p$ and $q$ iff there exists a positive constant $C_{p,q}$ such that
\begin{equation} \label{comp_def}
\comp(\varepsilon,d) \le C_{p,q}\, T(\varepsilon^{-1},d,p,q) \qquad \forall \varepsilon \in (0,\infty), \ \forall d \in \mathbb{N}.
\end{equation}
The  set of optimal parameters, $\mathcal{P}_{\text{opt}}$ is defined as
\begin{multline}
    \mathcal{P}_{\text{opt}} : = \{(p^*,q^*) : \eqref{comp_def} \text{ holds }\forall (p,q) \in [p^*,\infty) \times [q^*,\infty) \setminus (p^*,q^*) \\ 
    \eqref{comp_def} \text{ fails } \forall (p,q) \in [0,p^*] \times [0,q^*] \setminus (p^*,q^*) \}.
\end{multline}

>>>>

Certain kinds of common notions of tractability are
\begin{equation*}
    \begin{array}{r@{\qquad}c@{\qquad}c}
    & \comp(\varepsilon, d)\\
    \toprule
    \text{strong polynomial algebraic  tractability} & \Theta(\varepsilon^{-p})\\
    \text{polynomial algebraic tractability} & \Theta(\varepsilon^{-p}d^{q})\\
    \text{strong polynomial exponential tractability} &  \Theta\bigl([\log(1 + \varepsilon^{-1})]^p\bigr)\\
    \text{polynomial exponential tractability} & \Theta\bigl([\log(1 + \varepsilon^{-1})]^p  d^{q}\bigr)
    \end{array}
\end{equation*}
\fred{We need to fix these so that the tractability function does not vanish as $\varepsilon \to \infty$.}

@@

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow [1,\infty),
\]
A problem is $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ such that
\begin{equation} \label{comp_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon \in (0,\infty), \ \forall d \in \mathbb{N}.
\end{equation}

Note that the tractability function is bounded below away from zero even as $\varepsilon \to \infty$, so that the information based complexity is not expected to vanish under arbitrarily loose error requirements.  For this definition of tractability to make sense we assume that
\begin{equation}
	T \text{ is non-decreasing in all variables}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp).
\end{equation}
There is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsion \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in (1,\infty).
\end{equation}

If \eqref{comp_def} holds for some $\vp$, it clearly holds for larger, $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{comp_def} holds.
The  set of optimal parameters, $\mathcal{P}_{\text{opt}}$ is defined as
\begin{equation}
	\mathcal{P}_{\text{opt}} : = \{\vp^* : \eqref{comp_def} \text{ holds }\forall \vp \in [\vp,\boldsymbol{\infty}) \setminus \{\vp^*\} \text{ and }
	\eqref{comp_def} \text{ fails } \forall \vp \in [\vzero,\vp] \setminus \{\vp^*\} \}.
\end{equation}
In this article we allow $\vp$ to be a scalar or vector as the situation requires.

\subsection{Strong tractability}
